---
title: "Prediction Assignment"
author: "Ismail Che Ani"
date: "July 1, 2018"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

This Prediction Assigment Writeup is for Peer-graded assignment in Practical Machine Learning course, as part of the Specialization in Data Science. It need to publish in html format using RStudio with knitr function.  To predict the manner in which 6 participants performed some exercise as described in the quiz is the main goal of the project and will be applied to the 20 test cases available in the test data.It will also graded automatically after submit the quiz,

##Introduction

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

## 1. Prepare and Loading Dataset

```{r , warning=FALSE, message = FALSE, echo=TRUE}

# Load library
library(knitr)
library(caret)
library(rpart)
library(rpart.plot)
library(rattle)
library(randomForest)
library(corrplot)
set.seed(301) 

# Set the URLs
TrainDataURL <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
TestDataURL <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

# Download and clean the datasets
TrainingData <- read.csv(url(TrainDataURL), na.strings=c("NA","#DIV/0!",""))
TestingData <- read.csv(url(TestDataURL), na.strings=c("NA","#DIV/0!",""))

# Check for the datasets dimemsions
dim(TrainingData)
dim(TestingData)
```
## 2. Cleaning and Checking the Dataset

```{r , warning=FALSE, message = FALSE, echo=TRUE}
# Delete columns with missing values
TrainingData <-TrainingData[,colSums(is.na(TrainingData)) == 0]
TestingData <-TestingData[,colSums(is.na(TestingData)) == 0]

# Delete unused columns
TrainingData <-TrainingData[,-c(1:7)]
TestingData <-TestingData[,-c(1:7)]

# Check for the datasets dimemsions
dim(TrainingData)
dim(TestingData)

# Partitions training dataset
PartData <- createDataPartition(TrainingData$classe, p=0.7, list=FALSE)
TrainingSet <- TrainingData[PartData, ]
TestingSet <- TrainingData[-PartData, ]
dim(TrainingSet)
dim(TestingSet)
```
# 3.  Showing the frequency and examining the Training Dataset

```{r , warning=FALSE, message = FALSE, echo=TRUE}
# Plot the Frequency different levels graph

plot(TrainingSet$classe, col="blue", main="Different Levels of Frequency", xlab="classe", ylab="Frequency")
```

# 4. Building the Decision Tree Model

Using Decision Tree, we shouldn’t expect the accuracy to be high. In fact, anything around 80% would be acceptable.

```{r , warning=FALSE, message = FALSE, echo=TRUE}
set.seed(12345)
modFitDT <- rpart(classe ~ ., data = TrainingSet, method="class", control = rpart.control(method = "cv", number = 10))
fancyRpartPlot(modFitDT)
```
## 5. Building the Random Forest Model

Using random forest, the out of sample error should be small. The error will be estimated using the 40% testing sample. We should expect an error estimate of < 3%.

```{r , warning=FALSE, message = FALSE, echo=TRUE}
# Fitting model
model1 <- randomForest(classe ~., data=TrainingSet, method="class")

# Predicting
prediction1 <- predict(model1, TestingSet, Type="class")

# Testing
confusionMatrix(prediction1, TestingSet$classe)
```
```{r , warning=FALSE, message = FALSE, echo=TRUE}
modFitRF <- randomForest(classe ~ ., data = TrainingSet, method = "rf", importance = T, trControl = trainControl(method = "cv", classProbs=TRUE,savePredictions=TRUE,allowParallel=TRUE, number = 10))

plot(modFitRF)
```

# 6. Prediction using Generalized Boosted Model

```{r , warning=FALSE, message = FALSE, echo=TRUE}
# Fitting model
fitControl <- trainControl(method="repeatedcv", number=5, repeats=1)
model2 <- train(classe ~., data=TrainingSet, method="gbm", trControl=fitControl, verbose=FALSE)

# Predicting
prediction2 <- predict(model2, TestingSet)

# Testing
confusionMatrix(prediction2, TestingSet$classe)

```
The accuracy of Random Forest is 99.64, which is higher than that by Generalized Boosted. The Random Forest
model will be applied to predict the 20 quiz results.

```{r , warning=FALSE, message = FALSE, echo=TRUE}
predictTest <- predict(model1, TestingData)
predictTest
```
